---
title: 'tidytext: Text mining and analysis using tidy tools'
author: "Julia Silge and David Robinson"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r echo = FALSE}
library(knitr)
opts_chunk$set(message = FALSE, warning = FALSE)
```

The tidy data framework is... One popular field that the tidy data has not generally been applied to is text mining and analysis.

In particular, we provide tools to work with text in a "one-row-per-token-per-document" format as 

### The tidy text format

Tidy data as defined by Wickham 2014 fits the following rules:

* One row per observation
* One column per variable
* One table per type of observational unit

Wickham notes that datasets in this format "are easy to manipulate, model and visualize." These work . It is worth emphasizing that "tidy" is not a perjorative term; not all datasets.

We define the tidy text format as being one-token-per-document-per-row. For example, consider the opening lines to Jane Austen's "Pride and Prejudice", as provided by the janeaustenr package:

```{r}
library(dplyr)
library(janeaustenr)

prideprejudice[10:16]
```

We would define the tidy format of this text to look as follows: 

```{r}
library(tidytext)

pride_and_prejudice_td <- data_frame(book = "Pride and Prejudice",
                                     text = prideprejudice) %>%
  slice(-(1:6)) %>%
  mutate(line = row_number()) %>%
  unnest_tokens(word, text)

knitr::kable(head(pride_and_prejudice_td))
```

Unlike a document-term matrix, this preserves the ordering of terms. It also allows for extra information to be annotated in additional columns. For example

```{r}

```

#### Converting to tidy formats

Some text datasets come in the form of document-term (or term-document) matrices, such as the `DocumentTermMatrix` object from the `tm` package. One example is the `AssociatedPress` dataset provided by the topicmodels package. 

```{r}
data("AssociatedPress", package = "topicmodels")
AssociatedPress
```

If we want to work with this dataset.

```{r}
tidy(AssociatedPress)
```

This allows the use of 

### Converting from tidy text formats

The tidytext package doesn't expect a user to keep text data in a tidy form at all times during an analysis. Many existing text mining packages and algorithms use document-term matrices as their inputs and outputs, often for valid computational or user-interface reasons. Therefore, the tidytext package offers tools to convert to and from other structures into a tidy form.

For example, you can convert to a TermDocumentMatrix using tidytext's `cast_tdm` function:

```{r}
ap_tdm <- ap_filtered %>%
  cast_tdm(document, term, count)

ap_tdm
```

(The package also provide tools for converting to quanteda's dfm object with `cast_dfm`, or to a general `sparseMatrix` object with `cast_sparse`). This is useful, for example, for working with Latent Dirichlet Allocation from the `topicmodels` package, which takes a `DocumentTermMatrix` as input.

```{r}
library(topicmodels)
lda <- LDA(ap_tdm, k = 2)
```

After this modeling, an analyst may be interested in exploring the model using tidy tools. The broom package [cite] lays out a philosophy for this kind of model analysis, built on three verbs: tidy (component-level), augment (observation-level), and glance (model-level).

tidytext extends this by implementing these verbs for LDA objects:

```{r}
tidy(lda)
tidy(lda, matrix = "gamma")
```

Note that we can extract either beta (per-term-per-topic probablities) or gamma (per-document-per-topic probabilities) matrices in a tidy form. $\beta$ is useful for understanding what the topics each mean:

```{r}
# graph here
```

While gamma is useful for soft-clustering the documents:

```{r}

```

The tidytext package thus serves as an interface between tidy formats and . This allows text mining tools to be integrated into a pipeline based on dplyr, ggplot2, and other tools.
