<!-- README.md is generated from README.Rmd. Please edit that file -->
[![Build Status](https://travis-ci.org/juliasilge/tidytext.svg?branch=master)](https://travis-ci.org/juliasilge/tidytext)

tidytext: Text mining using dplyr, ggplot2, and other tidy tools
---------------

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

### Jane Austen's Novels Can Be So Tidy

```{r}
library(janeaustenr)
library(dplyr)
originalbooks <- bind_rows(
  data_frame(text = sensesensibility, book = "Sense & Sensibility"),
  data_frame(text = prideprejudice, book = "Pride & Prejudice"),
  data_frame(text = mansfieldpark, book = "Mansfield Park"),
  data_frame(text = emma, book = "Emma"),
  data_frame(text = northangerabbey, book = "Northanger Abbey"),
  data_frame(text = persuasion, book = "Persuasion")
) %>% mutate(book = factor(book, levels = unique(book)))
```

Where are the chapters?

```{r}
library(stringr)
originalbooks <- originalbooks %>%
  group_by(book) %>%
  mutate(linenumber = row_number(),
         chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]", 
                                                 ignore_case = TRUE)))) %>%
  ungroup()
```

Now we can use our new function for unnest and tokenizing. We can use the `tokenizers` package if installed, or else stick with `str_split`. The default tokenizing is for words, but other options include characters, sentences, lines, paragraphs, and a regex pattern. By default, `unnest_tokens` drops the original text.

```{r}
library(tidytext)
library(tokenizers)
books <- originalbooks %>%
  unnest_tokens(word, text)

books
```

We can remove stop words kept in a tidy data set in the `tidytext` package.

```{r}
books <- books %>%
  filter(!(word %in% stopwords$word))
```

Now, let's see what are the most common words in all the books as a whole.

```{r}
books %>% count(word, sort = TRUE) 
```

Sentiment analysis can be done as an inner join. Three sentiment lexicons are in the `tidytext` package in the `sentiment` dataset. Let's look at the words with a sadness score from the NRC lexicon. What are the most common sadness words in *Mansfield Park*?

```{r}
nrcsadness <- filter(sentiments, lexicon == "nrc" & sentiment == "sadness")
books %>% filter(book == "Mansfield Park") %>% 
  inner_join(nrcsadness) %>% count(word, sort = TRUE)
```

Or instead we could examine how sentiment changes changes during each novel. Let's find a sentiment score for each word using the Bing lexicon, then count the number of positive and negative words in defined sections of each novel.

```{r}
library(tidyr)
bing <- filter(sentiments, lexicon == "bing")
janeaustensentiment <- books %>% inner_join(bing) %>% 
  count(book, index = linenumber %/% 80, sentiment) %>% 
  spread(sentiment, n, fill = 0) %>% 
  mutate(sentiment = positive - negative)
```

Now we can plot these sentiment scores across the plot trajectory of each novel.

```{r, fig.width=9, fig.height=9, warning=FALSE}
library(ggplot2)

ggplot(janeaustensentiment, aes(index, sentiment, fill = book)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  facet_wrap(~book, ncol = 2, scales = "free_x")
```

### Most common positive and negative words

One advantage of having 

```{r}
bing_word_counts <- books %>%
  inner_join(bing) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()

bing_word_counts
```

```{r}
bing_word_counts %>%
  filter(n > 150) %>%
  mutate(n = ifelse(sentiment == "negative", -n, n)) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n, fill = sentiment)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ylab("Contribution to positivity/negativity")
```

This lets us spot an anomaly in the sentiment analysis- that the word "miss" is coded as negative.

### Wordclouds

We've seen that this works well with ggplot2. But having the words in a tidy format is useful for other plots as well.

For example, consider the wordcloud package.

```{r}
library(wordcloud)

books %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 50))
```

In other functions, such as `comparison.cloud`, you may need to turn it into a matrix with reshape2's acast:

```{r wordcloud}
library(reshape2)

books %>%
  inner_join(bing) %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>% 
  comparison.cloud(colors = c("#F8766D", "#00BFC4"),
                   max.words = 75)
```

### Combining With a Dictionary

Download a psych dictionary:

```{r}
RIDzipfile <- download.file("http://provalisresearch.com/Download/RID.ZIP", "RID.zip")
unzip("RID.zip")
RIDdict <- quanteda::dictionary(file = "RID.CAT", format = "wordstat")
file.remove("RID.zip", "RID.CAT", "RID.exc")
```

And tidy it:

```{r}
rid <- tidy(RIDdict, regex = TRUE) %>%
  rename(regex = word) %>%
  tbl_df()

rid
```

For now let's focus on the "secondary needs" type:

```{r}
secondary <- rid %>%
  filter(str_detect(category, "SECONDARY"))

secondary
```

Now we can use the [fuzzyjoin](http://github.com/dgrtwo/fuzzyjoin) package to join these with the Emma:

```{r. eval }
library(fuzzyjoin)

secondary_emma <- emma_words %>%
  regex_inner_join(secondary, by = c(word = "regex"))
```
