<!-- README.md is generated from README.Rmd. Please edit that file -->

tidytext: Text mining using dplyr, ggplot2, and other tidy tools
---------------

In progress at ROpenSci Unconf 2016.

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

Jane Austen's novels can be so tidy.

```{r}
library(janeaustenr)
library(dplyr)
originalbooks <- bind_rows(
  data_frame(text = sensesensibility, book = "Sense & Sensibility"),
  data_frame(text = prideprejudice, book = "Pride & Prejudice"),
  data_frame(text = mansfieldpark, book = "Mansfield Park"),
  data_frame(text = emma, book = "Emma"),
  data_frame(text = northangerabbey, book = "Northanger Abbey"),
  data_frame(text = persuasion, book = "Persuasion")
) %>% mutate(book = factor(book, levels = unique(book)))
```

Where are the chapters?

```{r}
library(stringr)
originalbooks <- originalbooks %>%
  group_by(book) %>%
  mutate(linenumber = row_number(),
         chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]", 
                                                 ignore_case = TRUE)))) %>%
  ungroup()
```

Now we can use our new function for unnest and tokenizing. We can use the `tokenizers` package if installed, or else stick with `str_split`. The default tokenizing is for words, but other options include characters, sentences, lines, paragraphs, and a regex pattern. By default, `unnest_tokens` drops the original text.

```{r}
library(tokenizers)
books <- originalbooks %>%
  unnest_tokens(word, text)

books
```

We can remove stop words kept in a tidy data set in the `tidytext` package.

```{r}
books <- books %>%
  filter(!(word %in% stopwords$word))
```

Now, let's see what are the most common words in all the books as a whole.

```{r}
books %>% count(word, sort = TRUE) 
```

Sentiment analysis can be done as an inner join. Three sentiment lexicons are in the `tidytext` package in the `sentiment` dataset. Let's look at negative words from the Bing lexicon. What are the most common negative words in *Mansfield Park*?

```{r}
negativebing <- filter(sentiments, lexicon == "bing" & sentiment == "negative")
books %>% filter(book == "Mansfield Park") %>% 
  inner_join(negativebing) %>% count(word, sort = TRUE)
```

Or instead we could examine how sentiment changes changes during each novel. Let's find a sentiment score for each word, then count the number of positive and negative words in defined sections of each novel.

```{r}
bing <- filter(sentiments, lexicon == "bing")
janeaustensentiment <- books %>% inner_join(bing) %>% 
  count(book, index = linenumber %/% 80, sentiment) %>% 
  spread(sentiment, n, fill = 0) %>% 
  mutate(sentiment = positive - negative)
```

Now we can plot these sentiment scores across the plot trajectory of each novel.

```{r, fig.width=12, fig.height=16}
library(ggplot2)
ggplot(janeaustensentiment, aes(index, sentiment, fill = book)) +
  geom_bar(stat = "identity") +
  facet_wrap(~book, ncol = 2, scales = "free_x") +
  labs(title = "Sentiment in Jane Austen's Novels",
       subtitle = "Tidy text analysis makes handling text easier for many tasks",
       y = "Sentiment Score",
       caption = "Texts sourced from Project Gutenberg") +
  theme(axis.title.x=element_blank()) +
  theme(axis.text.x=element_blank()) +
  theme(axis.ticks.x=element_blank()) +
  theme(legend.position="none")
```



### Combining with a dictionary

Download a psych dictionary:

```{r}
RIDzipfile <- download.file("http://provalisresearch.com/Download/RID.ZIP", "RID.zip")
unzip("RID.zip")
RIDdict <- dictionary(file = "RID.CAT", format = "wordstat")
file.remove("RID.zip", "RID.CAT", "RID.exc")
```

And tidy it:

```{r}
rid <- tidy(RIDdict, regex = TRUE) %>%
  rename(regex = word) %>%
  tbl_df()

rid
```

For now let's focus on the "secondary needs" type:

```{r}
secondary <- rid %>%
  filter(str_detect(category, "SECONDARY"))

secondary
```

Now we can use the [fuzzyjoin](TODO) package to join these with the Emma:

```{r. eval }
library(fuzzyjoin)

secondary_emma <- emma_words %>%
  regex_inner_join(secondary, by = c(word = "regex"))
```
