<!-- README.md is generated from README.Rmd. Please edit that file -->

tidytext: Text mining using dplyr, ggplot2, and other tidy tools
---------------

In progress at ROpenSci Unconf 2016.

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

```{r}
books <- bind_rows(
  data_frame(text = sensesensibility, book = "Sense & Sensibility"),
  data_frame(text = prideprejudice, book = "Pride & Prejudice"),
  data_frame(text = mansfieldpark, book = "Mansfield Park"),
  data_frame(text = emma, book = "Emma"),
  data_frame(text = northangerabbey, book = "Northanger Abbey"),
  data_frame(text = persuasion, book = "Persuasion")
)
```

Book analysis

```{r}
books <- books %>%
  group_by(book) %>%
  mutate(text = str_to_lower(text),
         line = row_number(),
         chapter = cumsum(str_detect(text, "^chapter [\\divxlc]"))) %>%
  ungroup()
```

Split by word:

```{r}
books %>%
  unnest(word = str_split(text, "[^a-z']")) %>%
  filter(word != "")
```

### Create word-per-row data frame

```{r}
library(dplyr)
library(tidyr)
library(stringr)
library(janeaustenr)

emma_words <- data_frame(word = emma) %>%
  mutate(line = row_number()) %>%
  unnest(word = str_split(str_to_lower(emma), "[^a-z']")) %>%
  filter(word != "")
```

We can remove stop words using `anti_join`:

```{r}
emma_words <- emma_words %>%
  filter(!(word %in% stopwords$word))
```

### Combining with a dictionary

Download a psych dictionary:

```{r}
RIDzipfile <- download.file("http://provalisresearch.com/Download/RID.ZIP", "RID.zip")
unzip("RID.zip")
RIDdict <- dictionary(file = "RID.CAT", format = "wordstat")
file.remove("RID.zip", "RID.CAT", "RID.exc")
```

And tidy it:

```{r}
rid <- tidy(RIDdict, regex = TRUE) %>%
  rename(regex = word) %>%
  tbl_df()

rid
```

For now let's focus on the "secondary needs" type:

```{r}
secondary <- rid %>%
  filter(str_detect(category, "SECONDARY"))

secondary
```

Now we can use the [fuzzyjoin](TODO) package to join these with the Emma:

```{r. eval }
library(fuzzyjoin)

secondary_emma <- emma_words %>%
  regex_inner_join(secondary, by = c(word = "regex"))
```
