<!-- README.md is generated from README.Rmd. Please edit that file -->

tidytext: Text mining using dplyr, ggplot2, and other tidy tools
---------------

**Authors:** [Julia Silge](http://juliasilge.com/), [David Robinson](http://varianceexplained.org/)<br/>
**License:** [MIT](https://opensource.org/licenses/MIT)

[![Build Status](https://travis-ci.org/juliasilge/tidytext.svg?branch=master)](https://travis-ci.org/juliasilge/tidytext)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/tidytext)](https://cran.r-project.org/package=tidytext)


```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-",
  message = FALSE
)
```

Using [tidy data principles](https://www.jstatsoft.org/article/view/v059i10) can make many text mining tasks easier, more effective, and consistent with tools already in wide use. Much of the infrastructure needed for text mining with tidy data frames already exists in packages like [dplyr](https://cran.r-project.org/package=dplyr), [broom](https://cran.r-project.org/package=broom), [tidyr](https://cran.r-project.org/package=tidyr) and [ggplot2](https://cran.r-project.org/package=ggplot2). In this package, we provide functions and supporting data sets to allow conversion of text to and from tidy formats, and to switch seamlessly between tidy tools and existing text mining packages.

### Installation

You can install this package from CRAN:

```{r, eval=FALSE}
install.packages("tidytext")
```


Or you can install the development version from Github with [devtools](https://github.com/hadley/devtools):

```{r, eval=FALSE}
library(devtools)
install_github("juliasilge/tidytext")
```

### Tidy text mining example: the `unnest_tokens` function

The novels of Jane Austen can be so tidy! Let's use the text of Jane Austen's 6 completed, published novels from the [janeaustenr](https://cran.r-project.org/package=janeaustenr) package, and bring them into a tidy format. janeaustenr provides them as a one-row-per-line format:

```{r}
library(janeaustenr)
library(dplyr)

original_books <- austen_books() %>%
  group_by(book) %>%
  mutate(linenumber = row_number()) %>%
  ungroup()

original_books
```

To work with this as a tidy dataset, we need to restructure it as **one-token-per-row** format. The `unnest_tokens` function is a way to convert a dataframe with a text column to be one-token-per-row:

```{r}
library(tidytext)
tidy_books <- original_books %>%
  unnest_tokens(word, text)

tidy_books
```

This function uses the [tokenizers package](https://github.com/lmullen/tokenizers) to separate each line into words. The default tokenizing is for words, but other options include characters, ngrams, sentences, lines, paragraphs, or separation around a regex pattern.

Now that the data is in one-word-per-row format, we can manipulate it with tidy tools like dplyr. We can remove stop words (kept in the tidytext dataset `stop_words`) with an `anti_join`.

```{r}
data("stop_words")
tidy_books <- tidy_books %>%
  anti_join(stop_words)
```

We can also use `count` to find the most common words in all the books as a whole.

```{r}
tidy_books %>%
  count(word, sort = TRUE) 
```

Sentiment analysis can be done as an inner join. Three sentiment lexicons are in the tidytext package in the `sentiment` dataset. Let's examine how sentiment changes changes during each novel. Let's find a sentiment score for each word using the Bing lexicon, then count the number of positive and negative words in defined sections of each novel.

```{r}
library(tidyr)
bing <- sentiments %>%
  filter(lexicon == "bing") %>%
  select(-score)

bing

janeaustensentiment <- tidy_books %>%
  inner_join(bing) %>% 
  count(book, index = linenumber %/% 80, sentiment) %>% 
  spread(sentiment, n, fill = 0) %>% 
  mutate(sentiment = positive - negative)

janeaustensentiment
```

Now we can plot these sentiment scores across the plot trajectory of each novel.

```{r, fig.width=9, fig.height=9, warning=FALSE}
library(ggplot2)

ggplot(janeaustensentiment, aes(index, sentiment, fill = book)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  facet_wrap(~book, ncol = 2, scales = "free_x")
```

### Another example: the `pair_count` function

Another function is `pair_count`, which counts pairs of items that occur together within a group. Let's count the words that occur together in the lines of *Pride and Prejudice*.

```{r}
pride_prejudice_words <- tidy_books %>%
  filter(book == "Pride & Prejudice")
pride_prejudice_words

word_cooccurences <- pride_prejudice_words %>%
  pair_count(linenumber, word, sort = TRUE)

word_cooccurences
```

This can be useful, for example, to plot a network of co-occuring words with the [igraph](http://igraph.org/) and [ggraph](https://github.com/thomasp85/ggraph) packages.

```{r, fig.width=9, fig.height=6}
library(igraph)
library(ggraph)

set.seed(2016)
word_cooccurences %>%
  filter(n >= 10) %>%
  graph_from_data_frame() %>%
  ggraph(layout = "fr") +
  geom_edge_link(aes(edge_alpha = n, edge_width = n)) +
  geom_node_point(color = "lightblue", size = 5) +
  geom_node_text(aes(label = name), vjust = 1.8) +
  theme_void()
```

For more examples of text mining using tidy data frames, see the tidytext vignette.

### Tidying document term matrices

Many existing text mining datasets are in the form of a DocumentTermMatrix class (from the tm package). For example, consider the corpus of 2246 Associated Press articles from the topicmodels dataset.

```{r}
library(tm)
data("AssociatedPress", package = "topicmodels")
AssociatedPress
```

If we want to analyze this with tidy tools, we need to transform it into a one-row-per-term data frame first with a `tidy` function. (For more on the tidy verb, [see the broom package](https://github.com/dgrtwo/broom)).

```{r}
tidy(AssociatedPress)
```

We could find the most negative documents:

```{r}
ap_sentiments <- tidy(AssociatedPress) %>%
  inner_join(bing, by = c(term = "word")) %>%
  count(document, sentiment, wt = count) %>%
  ungroup() %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative) %>%
  arrange(sentiment)
```

Or we can join the Austen and AP datasets and compare the frequencies of each word:

```{r}
comparison <- tidy(AssociatedPress) %>%
  count(word = term) %>%
  rename(AP = n) %>%
  inner_join(count(tidy_books, word)) %>%
  rename(Austen = n) %>%
  mutate(AP = AP / sum(AP),
         Austen = Austen / sum(Austen))

comparison

library(scales)
ggplot(comparison, aes(AP, Austen)) +
  geom_point() +
  geom_text(aes(label = word), check_overlap = TRUE,
            vjust = 1, hjust = 1) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  geom_abline(color = "red")
```

For more examples of working with objects from other text mining packages using tidy data principles, see the vignette on converting to and from document term matrices.

### Code of Conduct

This project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.
