---
title: "Introduction to tidytext"
author: "Julia Silge and David Robinson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to tidytext}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, echo=FALSE}
library(knitr)
opts_chunk$set(warning=FALSE)
```


### The Life-Changing Magic of Tidying Text

Using [tidy data principles](https://www.jstatsoft.org/article/view/v059i10) can make many text mining tasks easier, more effective, and  consistent with tools already in wide use. Much of the infrastructure needed for text mining with tidy data frames already exists in packages like dplyr, broom, and ggplot2; in this package, we go the rest of the way and provide tidying functions and supporting data sets to make analyzing text tidy.

### A few first tidy text mining examples

The novels of Jane Austen can be so tidy! Let's use the text of Jane Austen's 6 completed, published novels from the [janeaustenr package](https://github.com/juliasilge/janeaustenr) to explore some text mining. Let's mark each line number for the books in the original data frame, and find where the chapters are.


```{r}
library(tidytext)
library(janeaustenr)
library(dplyr)
library(stringr)
originalbooks <- austen_books() %>%
  group_by(book) %>%
  mutate(linenumber = row_number(),
         chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]",
                                                 ignore_case = TRUE)))) %>%
  ungroup()
```

Now we can use the tidytext function to unnest and tokenize; this function uses the [tokenizers package](https://github.com/lmullen/tokenizers). The default tokenizing is for words, but other options include characters, sentences, lines, paragraphs, and a regex pattern. By default, `unnest_tokens` drops the original text.

```{r}
originalbooks <- originalbooks %>%
  unnest_tokens(word, text)

originalbooks
```

We can remove stop words kept in a tidy data set in the tidytext package with an antijoin.

```{r}
books <- originalbooks %>%
  anti_join(tidytext::stopwords)
```

Now, let's see what the most common words are in all six novels as a whole.

```{r}
books %>%
  count(word, sort = TRUE)
```

Sentiment analysis can be done as an inner join. Three sentiment lexicons are in the tidytext package in the `sentiment` dataset. Let's look at the words with a joy score from the NRC lexicon. What are the most common joy words in *Emma*?

```{r}
nrcjoy <- sentiments %>%
  filter(lexicon == "nrc", sentiment == "joy")

books %>%
  filter(book == "Emma") %>%
  semi_join(nrcjoy) %>%
  count(word, sort = TRUE)
```

Or instead we could examine how sentiment changes during each novel. Let's find a sentiment score for each word using the Bing lexicon, then count the number of positive and negative words in defined sections of each novel.

```{r}
library(tidyr)
bing <- sentiments %>%
  filter(lexicon == "bing") %>%
  select(-score)

janeaustensentiment <- originalbooks %>%
  inner_join(bing) %>%
  count(book, index = linenumber %/% 80, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative)
```

Now we can plot these sentiment scores across the plot trajectory of each novel.

```{r, fig.width=7, fig.height=7, warning=FALSE}
library(ggplot2)

ggplot(janeaustensentiment, aes(index, sentiment, fill = book)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  facet_wrap(~book, ncol = 2, scales = "free_x")
```

### Most common positive and negative words

One advantage of having the data frame with both sentiment and word is that we can analyze word counts that contribute to each sentiment.

```{r}
bing_word_counts <- originalbooks %>%
  inner_join(bing) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()

bing_word_counts
```

This can be shown visually, and we can pipe straight into ggplot2 because of the way we are consistently using tools built for handling tidy data frames.

```{r, fig.width=7, fig.height=6}
bing_word_counts %>%
  filter(n > 150) %>%
  mutate(n = ifelse(sentiment == "negative", -n, n)) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n, fill = sentiment)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ylab("Contribution to sentiment")
```

This lets us spot an anomaly in the sentiment analysis; the word "miss" is coded as negative but it is used as a title for young, unmarried women in Jane Austen's works. If it were appropriate for our purposes, we could easily add "miss" to a custom stopwords list using `bind_rows`.

### Wordclouds

We've seen that this tidy text mining approach works well with ggplot2, but having our data in a tidy format is useful for other plots as well.

For example, consider the wordcloud package. Let's look at the most common words in Jane Austen's works as a whole again.

```{r, fig.height=6, fig.width=6}
library(wordcloud)

books %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100))
```

In other functions, such as `comparison.cloud`, you may need to turn it into a matrix with reshape2's `acast`. Let's do the sentiment analysis to tag positive and negative words using an inner join, then find the most common positive and negative words. Until the step where we need to send the data to `comparison.cloud`, this can all be done with joins, piping, and dplyr because our data is in tidy format.

```{r wordcloud, fig.height=5, fig.width=5}
library(reshape2)

books %>%
  inner_join(bing) %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("#F8766D", "#00BFC4"),
                   max.words = 100)
```

### Looking at units beyond just words

Lots of useful work can be done by tokenizing at the word level, but sometimes it is useful or necessary to look at different units of text. For example, some sentiment analysis algorithms look beyond only unigrams (i.e. single words) to try to understand the sentiment of a sentence as a whole. These algorithms try to understand that

> I am not having a good day.

is a sad sentence, not a happy one, because of negation. The [Stanford CoreNLP](http://stanfordnlp.github.io/CoreNLP/) tools and the [sentimentr R package](https://github.com/trinker/sentimentr) (currently available on Github but not CRAN) are examples of such sentiment analysis algorithms. For these, we may want to tokenize text into sentences.

```{r}
austen_sentences <- austen_books() %>% 
  group_by(book) %>% 
  unnest_tokens(sentence, text, token = "sentences") %>% 
  ungroup()
```

Let's look at just one.

```{r}
austen_sentences$sentence[39]
```

The sentence tokenizing does seem to have a bit of trouble with UTF-8 encoded text, especially with sections of dialogue; it does much better with punctuation in ASCII.

Near the beginning of this vignette, we used a regex to find where all the chapters were in Austen's novels. We can use tidy text analysis to ask questions such as what are the most negative chapters in each of Jane Austen's novels? First, let's get the list of negative words from the Bing lexicon. Second, let's make a dataframe of how many words are in each chapter so we can normalize for the length of chapters. Then, let's find the number of negative words in each chapter and divide by the total words in each chapter. Which chapter has the highest proportion of negative words?

```{r}
bingnegative <- sentiments %>%
  filter(lexicon == "bing", sentiment == "negative")

wordcounts <- originalbooks %>%
  group_by(book, chapter) %>%
  summarize(words = n())

originalbooks %>%
  semi_join(bingnegative) %>%
  group_by(book, chapter) %>%
  summarize(negativewords = n()) %>%
  left_join(wordcounts, by = c("book", "chapter")) %>%
  mutate(ratio = negativewords/words) %>%
  filter(chapter != 0) %>%
  top_n(1)
```

These are the chapters with the most sad words in each book, normalized for number of words in the chapter. What is happening in these chapters? In Chapter 43 of *Sense and Sensibility* Marianne is seriously ill, near death, and in Chapter 34 of *Pride and Prejudice* Mr. Darcy proposes for the first time (so badly!). Chapter 46 of *Mansfield Park* is almost the end, when everyone learns of Henry's scandalous adultery, Chapter 15 of *Emma* is when horrifying Mr. Elton proposes, and in Chapter 21 of *Northanger Abbey* Catherine is deep in her Gothic faux fantasy of murder/etc. Chapter 4 of *Persuasion* is when the reader gets the full flashback of Anne refusing Captain Wentworth and how sad she was and what a terrible mistake she realized it to be.
